Stability of variational quantum dynamics with Gaussian approximants

### 2013-10-31 paper.tex
writing

Consider a linear combination of kets,
$$|ψ〉=c₁|a₁〉+c₂|a₂〉+⋯+c_R|a_R〉.$$
This can be written as a matrix product $|ψ〉=|A〉c$, where $c=\pmatrix{c₁&⋯&c_R}^{\rm T}$ is a column vector of coefficients, and $|A〉=\pmatrix{|a₁〉&⋯&|a_R〉}$ is a row vector of kets.  Alternatively, the kets $|a_r〉$ can be regarded as infinite dimensional columns, and $|A〉$ as a matrix, representing a linear operator from ${\bf C}^R$ into ket space.  This operator has an adjoint, $〈A|=\pmatrix{〈a₁|&⋯&〈a_R|}^{\rm T}$, that maps kets back to column vectors.  Under the rules of matrix multiplication, the product
$$〈A|A〉=\pmatrix{〈a₁|a₁〉&&〈a₁|a_R〉\cr &\ddots&\cr 〈a_R|a₁〉&&〈a_R|a_R〉}$$
is an operator on ${\bf C}^R$, and $|A〉〈A|=|a₁〉〈a₁|+⋯+|a_R〉〈a_R|$ is a Hilbert space operator.  Other matrix products work too; provided the $|a_r〉$ are linearly independent, $|A〉$ will have a singular value decomposition $|A〉=|U〉SV†$, where $|U〉$ is a row of $R$ kets, $S$ is an $R×R$ diagonal matrix of complex numbers, and $V$ is an orthogonal $R×R$ complex matrix%
\footnote{${}^1$}{In fact, $|A〉$ has an SVD regardless, but $|U〉$ and $S$ are smaller.  When the $|a_r〉$ are linearly dependent, they span a space of dimension less than $R$, and there is no way to choose in which directions it should be extended to give $R$ orthonormal kets.  In this case, $|U〉$ will be a row of ${\rm rank}|A〉$ kets, and $S$ will be ${\rm rank}|A〉×R$.}.
In this paper, things written as kets have a finite number of infinite dimensional columns, and bras are the other way round.

@@

The approximant studied in this paper has the form
$$|ψ(φ₁,α₁,⋯,φ_R,α_R)〉=∑_{r=1}^R e^{φ_r+α_ra†}|0〉
	=\pmatrix{|α₁〉&⋯&|α_R〉}\pmatrix{e^{φ₁+½|α₁|²}\cr\vdots\cr e^{φ_R+½|α_R|²}}
	=|A〉d,\tag\Ad$$
where $A=\pmatrix{α₁&⋯&α_R}^{\rm T}$ is a vector of complex amplitudes, $|α_r〉$ is a coherent state, and $φ_r$ is a complex weight.  The form on the left is more complicated than the $|A〉d$ one, but it is an analytic function of the parameters, which is convenient for derivatives and optimisation.  Wavelet theorists have studyed the conditions on the set of ampltiudes $A$ that allow any state to be expanded in this form, and the stability of expansions as the density of ampltiudes decreases towards the minimum for completeness.  This will be discussed later.

In the limit of large $R$, this sum approximates an integral over the amplitude $α$, and any state can be approximated arbitrarily well by this form.  The continuous phase-space expansion is the topic of this and the next section.  Many problems in quantum optics have been solved by a family of Monte-Carlo methods, using similar expansions, where the amplitudes $α_r$ are propagated independently.  This gives some grounds to hope that the states of interesting many-particle systems can be expanded with a reasonable number of components $R$.

Expansion over coherent states is a linear operator from functions of a complex variable to quantum states,
$$\G f=π^{-½}\int |α〉e^{-½|α|²}f(α*)\,d²α.$$
The notation $\int f(α)\,d²α$ means $\int_{-∞}^{∞}\int_{-∞}^{∞} f(x+iy)\,dx\,dy$.  The function $f$ is such that $e^{-½|α|²}f(α)$ is square-integrable.  Bargmann\cite{pam-14-187} showed that these functions form a Hilbert space under the inner product $〈g,f〉=\int e^{-|α|²}g*(α)f(α)\,d²α$; this paper refers to it as the space of phase space distributions, or just distributions.  When norms of functions are written below, the norm is the one derived from this inner product.  In this paper, $\G f$ denotes the ket that results from applying the linear operator $\G$ to the phase space distribution $f$; application and composition of linear operators is shown by juxtaposition, and functions whose values are quantum states are written as kets, usually with a greek letter.

@@

\midinsert \XeTeXpicfile tysv.pdf width \hsize
Figure 3 \endinsert

>>>>

Consider a linear combination of kets,
$$|ψ〉=c₁|a₁〉+c₂|a₂〉+⋯+c_R|a_R〉.$$
This can be written as a matrix product $|ψ〉=|A〉c$, where $c=\pmatrix{c₁&⋯&c_R}^{\rm T}$ is a column vector of coefficients, and $|A〉=\pmatrix{|a₁〉&⋯&|a_R〉}$ is a row vector of kets.  Alternatively, the kets $|a_r〉$ can be regarded as infinite dimensional columns, and $|A〉$ as a matrix, representing a linear operator from ${\bf C}^R$ into ket space.  This operator has an adjoint, $〈A|=\pmatrix{〈a₁|&⋯&〈a_R|}^{\rm T}$, that maps kets back to column vectors.  Under the rules of matrix multiplication, the product
$$〈A|A〉=\pmatrix{〈a₁|a₁〉&&〈a₁|a_R〉\cr &\ddots&\cr 〈a_R|a₁〉&&〈a_R|a_R〉}$$
is an operator on ${\bf C}^R$, and $|A〉〈A|=|a₁〉〈a₁|+⋯+|a_R〉〈a_R|$ is a Hilbert space operator.  Other matrix products work too; provided the $|a_r〉$ are linearly independent, $|A〉$ will have a singular value decomposition $|A〉=|U〉SV†$, where $|U〉$ is a row of $R$ kets, $S$ is an $R×R$ diagonal matrix of complex numbers, and $V$ is an orthogonal $R×R$ complex matrix%
\footnote{${}^1$}{In fact, $|A〉$ has an SVD regardless, but $|U〉$ and $S$ are smaller.  When the $|a_r〉$ are linearly dependent, they span a space of dimension less than $R$, and there is no way to choose in which directions it should be extended to give $R$ orthonormal kets.  In this case, $|U〉$ will be a row of ${\rm rank}|A〉$ kets, and $S$ will be ${\rm rank}|A〉×R$.}.
In this paper, things written as kets are linear mappings from other things to quantum states.  The other things might be finite vectors of complex numbers, in which case the ket can be thought of as a matrix, or they might be functions from an infinite dimensional space, in which case this is less useful.  Bras are the other way round: they map states to things other than states.  Ordinary bras and kets are a special case.

@@

The approximant studied in this paper has the form
$$|ψ(φ₁,α₁,⋯,φ_R,α_R)〉=∑_{r=1}^R e^{φ_r+α_ra†}|0〉
	=\pmatrix{|α₁〉&⋯&|α_R〉}\pmatrix{e^{φ₁+½|α₁|²}\cr\vdots\cr e^{φ_R+½|α_R|²}}
	=|A〉d,\tag\Ad$$
where $A=\pmatrix{α₁&⋯&α_R}^{\rm T}$ is a vector of complex amplitudes, $|α_r〉$ is a coherent state, and $φ_r$ is a complex weight.  The form on the left is more complicated than the $|A〉d$ one, but it is an analytic function of the parameters, which is convenient for derivatives and optimisation.  Wavelet theorists have studied the conditions on the set of ampltiudes $A$ that allow any state to be expanded in this form, and the stability of expansions as the density of ampltiudes decreases towards the minimum for completeness.  This will be discussed later.

In the limit of large $R$, this sum approximates an integral over the amplitude $α$; any state can be approximated arbitrarily well by this form.  The continuous phase-space expansion is the topic of this and the next section.  Many problems in quantum optics have been solved by a family of Monte-Carlo methods, using similar expansions, where the amplitudes $α_r$ are propagated independently.  This gives some grounds to hope that the states of interesting many-particle systems can be expanded with a reasonable number of components $R$.

Expansion over coherent states is a linear operator from functions $f$ of a complex variable to quantum states $\G f$,
$$\G f=π^{-½}\int |α〉e^{-½|α|²}f(α*)\,d²α.$$
The notation $\int f(α)\,d²α$ means $\int_{-∞}^{∞}\int_{-∞}^{∞} f(x+iy)\,dx\,dy$.  The function $f$ is such that $e^{-½|α|²}f(α)$ is square-integrable.  Bargmann\cite{pam-14-187} showed that these functions form a Hilbert space under the inner product $〈g,f〉=\int e^{-|α|²}g*(α)f(α)\,d²α$; this paper refers to it as the space of phase space distributions, or just distributions.  When norms of functions are written below, the norm is the one derived from this inner product.

@@

\midinsert %\XeTeXpicfile tysv.pdf width \hsize
Figure 3 \endinsert
